---
title: "The Secret of the Winnest Team"
author: "Tony Wen"
date: "4/7/2022"
output: html_document
runtime: shiny
---

##### package deposit
```{r warning = F, message = F}
library(tidyverse)
library(ggplot2); library(ggpubr)
library(qacBase); library(qacDR)
library(Rtsne); library(cluster)

library(ggalluvial)
library(igraph);library(networkD3) # (interactive) network

library(kableExtra)

library(fmsb) # radar plot

library(shiny)
```

### Data Collection
+ Placeholder for final report
+ will have a summary with link to python api requests

### Preliminary Data Analysis and Management

```{r warning = F, message = F}
# load in collected data
player_stats = read.csv("winnestTeam_stats.csv")
lineup_stats = read.csv("lineup_stats.csv")
assist_net = read.csv("AssistNetwork.csv")
```

##### Check Missing Values
```{r}
contents(player_stats)$overall %>% mutate(pct_miss = as.numeric(gsub("%", "", pct_miss))) %>% arrange(desc(pct_miss)) %>%
  select(pct_miss) %>% kbl() %>%
  kable_paper("hover") %>%
  scroll_box(height = "250px")
```

> Some of the variables are extremely specific and does not really help in context

+ Period`x`Fouls`y`Minutes
+ HeaveMakes
+ HeaveAttempts
+ Clear.Path.Fouls
+ OffensiveGoaltends
+ PeriodOTFouls`X`Minutes

> Some of the NAs indicates 0s

+ ``.Trips
+ Drawn
+ Blocked


##### Factor Analysis after setting NAs to 0s
```{r message = F, warning = F, echo = F}
# try to replace NAs with 0s
# and get rid of time-dependent fouls
player_stats %>% replace(is.na(.), 0) %>% select(-contains("Period")) -> player_stats_mod

player_stats_mod %>% .[11:ncol(player_stats_mod)] %>%
  scree_plot(method = "pa")
```

```{r results = F}
# FA for a factor of 4
player_stats_model = player_stats_mod  %>% .[11:ncol(player_stats_mod)] %>% FA(nfactors = 4, rotate = "promax")
```

```{r}
# check out the factors
player_stats_model %>% plot()

#
player_stats_factor = score(player_stats %>% replace(is.na(.), 0),player_stats_model)
```

check out the content of a factor
```{r}
player_stats_model$Structure %>% 
  mutate(prime_factor = case_when(pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F1) ~ "F1",
                                  pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F2) ~ "F2",
                                  pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F3) ~ "F3",
                                  pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F4) ~ "F4")) %>% 
  filter(prime_factor == "F4") %>% arrange(desc(F4))
```

convert Factor structure into a network for visualization
```{r warning = F, message = F, fig.height = 20, fig.width = 20}
player_stats_model$Structure %>% add_rownames(var = "From") %>%
  mutate(prime_factor = case_when(pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F1) ~ "F1",
                                  pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F2) ~ "F2",
                                  pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F3) ~ "F3",
                                  pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F4) ~ "F4")) %>%
  gather("To", "Weight", 2:5)%>%
  #reorder
  select(From, To, Weight, prime_factor) %>%
  mutate(edgeCol = case_when(To == "F1" ~ "red",
                             To == "F2" ~ "orange",
                             To == "F3" ~ "green",
                             To == "F4" ~ "blue"))-> pstats_from_to

pstats_from_to %>% select("From") %>% unique() %>%
  left_join(pstats_from_to, by = "From") %>%
  mutate(color = case_when(prime_factor == "F1" ~ "red",
                           prime_factor == "F2" ~ "orange",
                           prime_factor == "F3" ~ "green",
                           prime_factor == "F4" ~ "blue"),
         size = ifelse(prime_factor %in% c("F1", "F2", "F3", "F4"), 10, 5)) %>%
  select(From, color, size, prime_factor)%>% 
  rbind(tibble(From = c("F1", "F2", "F3", "F4"),
               color = c("red", "orange", "green", "blue"),
               size = c(10, 10, 10, 10),
               prime_factor = c("F1", "F2", "F3", "F4"))) %>% unique() -> pstats_node 

# plot only a network for selected Factor
plot(graph_from_data_frame(pstats_from_to %>% filter(prime_factor == "F1"),
                           directed = F,
                           pstats_node %>% filter(color == "red"|From %in% c("F1", "F2", "F3", "F4"))),
     edge.width = (pstats_from_to %>% filter(prime_factor == "F1"))$Weight*5,
     edge.color = (pstats_from_to %>% filter(prime_factor == "F1"))$edgeCol)
```

convert factor structure into an alluvial plot
```{r warning = F, message = F}
# weight controls the direction of the correlation
pstats_from_to %>% filter(prime_factor == "F4", Weight > 0) %>%
  ggplot(aes(axis1 = From, axis2 = To, y = Weight))+
  geom_flow(aes(fill = To))+
  geom_stratum()+
  geom_text(stat = "stratum", infer.label = T)+
  scale_x_discrete(limits = c("Stats", "Factors"))
```

build a shiny app for all three visual breakdown of the factors
```{r}
ui = fluidPage(
  titlePanel(textOutput("title")),
  sidebarLayout(
    sidebarPanel(
      checkboxGroupInput(
        inputId = "FactorInvolved",
        label = "Select the Factors Involved (Best for Network)",
        choices = c("F1", "F2", "F3", "F4"),
        selected = "F1"
      ),
      radioButtons(
        inputId = "Polarity",
        label = "Select the Direction of the Correlation (Alluvial Only)",
        choices = c("Positive", "Negative"),
        selected = "Positive"
      )
    ),
    
    mainPanel(
      tabsetPanel(
        tabPanel("Network", plotOutput(outputId = "network")),
        tabPanel("Alluvial", plotOutput(outputId = "alluvial")),
        tabPanel("Table", tableOutput(outputId = "FactorComp"))
      )
    )
  )
)


server = function(input, output){
  
  output$title = renderText("Factor Breakdown")
  
  output$FactorComp = function()(
    {
      player_stats_model$Structure %>% 
        mutate(prime_factor = case_when(pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F1) ~ "F1",
                                        pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F2) ~ "F2",
                                        pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F3) ~ "F3",
                                        pmax(abs(F1), abs(F2), abs(F3), abs(F4)) == abs(F4) ~ "F4")) %>% 
        filter(prime_factor %in% input$FactorInvolved) %>% arrange(desc(input$FactorInvolved[1])) -> result
      
      result %>% kable("html", align = "ccccc") %>%
        row_spec(seq(1,nrow(result), 2), 
                 background = "gray",
                 color = "white") %>%
        kable_paper("hover") %>%
        scroll_box(height = "500px")
    }
  )
  
  output$network = renderPlot(
    {
      plot(graph_from_data_frame(pstats_from_to %>% filter(prime_factor %in% input$FactorInvolved),
                                 directed = F,
                                 pstats_node %>% filter(prime_factor %in% input$FactorInvolved|From %in% c("F1", "F2", "F3", "F4"))),
           edge.width = (pstats_from_to %>% filter(prime_factor %in% input$FactorInvolved))$Weight,
           edge.color = (pstats_from_to %>% filter(prime_factor %in% input$FactorInvolved))$edgeCol)
    }, height = 1000, width = 1000
  )
  
  output$alluvial = renderPlot(
    {
      if (input$Polarity == "Positive"){
        pstats_from_to %>% filter(prime_factor %in% input$FactorInvolved, 
                                  Weight > 0) %>%
          ggplot(aes(axis1 = From, axis2 = To, y = Weight))+
          geom_flow(aes(fill = To))+
          geom_stratum()+
          geom_text(stat = "stratum", infer.label = T)+
          scale_x_discrete(limits = c("Stats", "Factors"))}
      else{
        pstats_from_to %>% filter(prime_factor %in% input$FactorInvolved, 
                                  Weight < 0) %>%
          ggplot(aes(axis1 = From, axis2 = To, y = Weight))+
          geom_flow(aes(fill = To))+
          geom_stratum()+
          geom_text(stat = "stratum", infer.label = T)+
          scale_x_discrete(limits = c("Stats", "Factors"))
      }
    }, height = 800, width = 1000
  )
}

shinyApp(ui, server, options = list(height = 1000))
```

> Based on the graphics and the variables in each factors, it seems that:
<br>+ F1 = Paint Master(close-to-mid range scoring + Defense)
<br>+ F2 = Long Range General (3pt+Assist)
<br>+ F3 = 3pt Conversion/Efficiency
<br>+ F4 = At Rim Conversion/Efficiency


### Check out how players fit into different Factors
```{r}
player_stats_factor %>% select(Year, Name, F1:F4) %>%
  mutate(player = paste0(Year,"-", Name)) %>% select(-Year, -Name) %>%
  column_to_rownames(var = "player") %>% scale() %>% as.data.frame() -> player_factor_std

qacDR::wss_plot(player_factor_std) # maybe 5?
```

```{r}
set.seed(1234)
fit.km = kmeans(player_factor_std, 5, nstart = 25)

# put cluster membership back in player_factor_std
player_factor_std = player_factor_std %>% mutate(cluster = factor(fit.km$cluster))

# profile plot
qacDR::profile_plot(player_factor_std, type = "bar")
```

```{r message = F, warning = F, fig.width = 16, fig.height =  16}
set.seed(1234)

score(player_factor_std, PCA(player_factor_std[1:4], nfactors = 2)) %>%
  ggplot(aes(x = PC1, y = PC2, color = cluster))+
  geom_point()+
  ggrepel::geom_text_repel(label = rownames(player_factor_std), size = 2)

# come back and annotate this ^^^

# tsne projection
#(cluster::daisy(player_factor_std[1:4], metric = "euclidean") %>% Rtsne(is_distance = T))$Y %>% data.frame() %>%
#  setNames(c("Dim1", "Dim2")) %>% 
#  mutate(cluster = factor(player_factor_std$cluster),
#         name = rownames(player_factor_std)) %>%
#  ggplot(aes(x = Dim1, y = Dim2, color = cluster))+
#  geom_point()+
#  geom_text_repel(label = rownames(player_factor_std))
```

```{r}
colors_border=c( rgb(0.2,0.5,0.5,0.9), rgb(0.8,0.2,0.5,0.9) , rgb(0.7,0.5,0.1,0.9) )
colors_in=c( rgb(0.2,0.5,0.5,0.4), rgb(0.8,0.2,0.5,0.4) , rgb(0.7,0.5,0.1,0.4) )

rbind(rep(2.5,5) , rep(-2.5,5), 
      (player_factor_std %>% rownames_to_column("player") %>% 
         filter(player %in% c("2004-Allen Iverson",
                              "2017-James Harden",
                              "2006-Dirk Nowitzki")) %>% .[2:5])) %>% 
  radarchart(pcol = colors_border, pfcol = colors_in)

# ^^^ work-in-progress shiny for above ^^^
```


### NEXTUP: Aggregate Factors based on Team Lineups

### NEXTUP: Incorporate Assist Net Somehow?





